Best Practices for Alchemical Free Energy Calculations : v\versionnumber

### 2018-04-11 manuscript.tex

\author[1*]{John D. Chodera}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[4*]{Conor Parks}
\author[5*]{Julia E. Rice}
\author[6*]{Michael Shirts}
\author[7*]{Bryce Allen}
\author[1]{Levi N. Naden}
\author[1,9]{Andrea Rizzi}
\affil[1]{Computational and Systems Biology Program, Memorial Sloan Kettering Cancer Center, New York NY 10065}
\affil[2]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine}
\affil[3]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[4]{I don't know my affiliation}
\affil[5]{I don't know my affiliation}
\affil[8]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}

@@

\begin{enumerate}
\item Determine whether you need to \textbf{handle multiple binding modes}: (Mobley)
\begin{itemize}
\item In absolute: Confine-and-release, BLUES
\item In relative (ugh!!)
\end{itemize}
\todo[inline, color={red!40}]{LNN, BR: write stopping condition section}
\item Determine \textbf{stopping conditions}
Uncertainty-directed stopping criteria can ensure target uncertainty is achieved

@@

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BR: uncertainty estimation section}
\begin{enumerate}
\item Detecting boundary between equilibrated and production regions (Chodera: \url{http://dx.doi.org/10.1021/acs.jctc.5b00784})
\item Decorrelating samples for analysis
\begin{itemize}
\item Subsample different lambdas based on correlation times
\item Ensure all simulations at least 50x correlation time
\end{itemize}
\item Examining output data for common problems with discussions of what exactly to plot or look at; examples of typical curves for dV/dlambda and free energy versus lambda, for example
\begin{itemize}
\item Make sure ligand doesn’t tumble out of binding site (Mey has observed this)
\item Significant discrepancies between different free energy estimators (TI, BAR, MBAR)
\item Poor replica mixing (for replica-exchange)
\item Correlation time as a function of lambda as it would be expected to be a smooth
\item Dependence on initial conformation
\item Torsional analysis: Is it stuck in specific states? Only very rarely transitions?
\item More “usual suspects”
\end{itemize}
\item Estimators for free energies
\begin{itemize}
\item MBAR recommended if all energy differences are available
\item BAR just as good for highly optimized lambda values
\item TI should be roughly concordant, but quadrature error hard to quantify
\item Other variants useful in special circumstances (e.g. Z. Tan stochastic version)
\end{itemize}
\item Computing and reporting uncertainties on free energies
correlated bootstrap v. timeseries analysis
\item Other considerations for many transformations
Cycle closure error

>>>>

\author[1*]{John D. Chodera}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[4*]{Conor Parks}
\author[5*]{Julia E. Rice}
\author[6*]{Michael Shirts}
\author[7]{Bryce K. Allen}
\author[1]{Levi N. Naden}
\author[1,9]{Andrea Rizzi}
\affil[1]{Computational and Systems Biology Program, Memorial Sloan Kettering Cancer Center, New York NY 10065}
\affil[2]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine}
\affil[3]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[4]{I don't know my affiliation}
\affil[5]{I don't know my affiliation}
\affil[7]{Silicon Therapeutics, Boston, MA, USA}
\affil[8]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}

@@

\begin{enumerate}
\item Determine whether you need to \textbf{handle multiple binding modes}: (Mobley)
\begin{itemize}
\item In absolute: Confine-and-release, BLUES
\item In relative (ugh!!)
\end{itemize}
\todo[inline, color={red!40}]{LNN, BA: write stopping condition section}
\item Determine \textbf{stopping conditions}
Uncertainty-directed stopping criteria can ensure target uncertainty is achieved

@@

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}
\begin{enumerate}
\item Detecting boundary between equilibrated and production regions (Chodera: \url{http://dx.doi.org/10.1021/acs.jctc.5b00784})
\item Decorrelating samples for analysis
\begin{itemize}
\item Subsample different lambdas based on correlation times
\item Ensure all simulations at least 50x correlation time
\end{itemize}
\item Examining output data for common problems with discussions of what exactly to plot or look at; examples of typical curves for dV/dlambda and free energy versus lambda, for example
\begin{itemize}
\item Make sure ligand doesn’t tumble out of binding site (Mey has observed this)
\item Significant discrepancies between different free energy estimators (TI, BAR, MBAR)
\item Poor replica mixing (for replica-exchange)
\item Correlation time as a function of lambda as it would be expected to be a smooth
\item Dependence on initial conformation
\item Torsional analysis: Is it stuck in specific states? Only very rarely transitions?
\item More “usual suspects”
\end{itemize}
\item Estimators for free energies
\begin{itemize}
\item MBAR recommended if all energy differences are available
\item BAR just as good for highly optimized lambda values
\item TI should be roughly concordant, but quadrature error hard to quantify
\item Other variants useful in special circumstances (e.g. Z. Tan stochastic version)
\end{itemize}
\item Computing and reporting uncertainties on free energies
correlated bootstrap v. timeseries analysis
\begin{itemize}
\item Quantifying standard error in dG estimate
\item Varience in dG estimate using multiple methods (TI, DEXP, IEXP, BAR, MBAR, GDEL, etc.)
\item Agreement in dG estimate when repeating calculations with different parameters (random seeds, initial configurations, forcefield, etc.)
\item Calculating differences in free energy change as a function of time
\begin{itemize}
	\item Starting from beginning or end of simulation
	\item Significance of differences in midpoint estimate (High middle error: high uncertainty)
\end{itemize}
\item Ensemble method to combine uncertainties into interpretable weighted metric
\begin{itemize}
	\item Simple normalized metric to determine confidence in calculation
	\item Easily interpreted by chemist/biologist when prioritizing new chemistries
\end{itemize}
\end{itemize}
\item Other considerations for many transformations
Cycle closure error
